Cooperative Kernels: Safe Multitasking for Blocking Algorithms on GPUs

### 2016-11-15 write-ups/PLDI/paper.tex
Changes with Hugues.
worked on section 4

There is scope for combining the approaches: if a
cooperative kernel is insufficiently cooperative, i.e.\ it does not
invoke $\offerkill$ enough, then hardware support for preemption could
be invoked to prevent the kernel from denying other tasks access to
the GPU indefinitely.

\paragraph{Optimizing GPU scheduling}
%
A number of works have considered how best to schedule dynamic
workloads on GPUs~\cite{DBLP:journals/tog/SteinbergerKKHKS12,DBLP:conf/ics/WuCLSV15,DBLP:conf/usenix/KatoLRI11,DBLP:journals/tog/SteinbergerKBKDS14,DBLP:journals/rts/ElliottA12,DBLP:conf/sosp/RossbachCSRW11}.  Among these, the Whippletree approach employs a persistent
megakernel to schedule multiple, dynamic tasks in a manner that aims
to best utilise GPU
resources~\cite{DBLP:journals/tog/SteinbergerKBKDS14}.  The TimeGraph
approach similarly aims to optimise scheduling of competing workloads,
in the context of OpenGL, using GPU-CPU synchronisation so that the
CPU can monitor and control the work that is scheduled on the
GPU~\cite{DBLP:conf/usenix/KatoLRI11}.  To best utilise the compute
units of a GPU for a particular kernel, SM-centric program
transformations have been proposed, which aim to direct work to
specific compute units, rather than to specific workgroups, based on
the observation that in practice multiple workgroups may share the
same compute unit, so that said workgrops might be in a good position
to exploit memory caches associated with that compute
unit~\cite{DBLP:conf/ics/WuCLSV15}.  None of these approaches tackle
the problem of \emph{fair} scheduling on GPUs that our work addresses, thus they do not aid in safe deployment of blocking irregular algorithms.

\paragraph{Dynamic parallelism} Recent version of CUDA and OpenCL allow a kernel to spawn further kernels~\cite{...}.
This \emph{dynamic parallelism} can be exploited to implement a GPU-based scheduler: an initial scheduler kernel uses the feature to spawn additional kernels as required~\cite{ }.

 is launched

@@

\cite{TPO10}

>>>>

\ADComment{Rewrite:} There is scope for combining the approaches by using cooperative scheduling...

\paragraph{Optimizing GPU scheduling}
%
A number of works have considered how best to schedulel dynamic
workloads on GPUs.  The Whippletree approach employs a persistent
megakernel to schedule multiple, dynamic tasks in a manner that aims
to best utilise GPU
resources~\cite{DBLP:journals/tog/SteinbergerKBKDS14}.  The TimeGraph
approach similarly aims to optimise scheduling of competing workloads,
in the context of OpenGL, using GPU-CPU synchronisation so that the
CPU can monitor and control the work that is scheduled on the
GPU~\cite{DBLP:conf/usenix/KatoLRI11}.  To best utilise the compute
units of a GPU for a particular kernel, SM-centric program
transformations have been proposed, which aim to direct work to
specific compute units, rather than to specific workgroups, based on
the observation that in practice multiple workgroups may share the
same compute unit, so that said workgrops might be in a good position
to exploit memory caches associated with that compute
unit~\cite{DBLP:conf/ics/WuCLSV15}.  None of these approaches tackle
the problem of \emph{fair} scheduling on GPUs that our work addresses, thus they do not aid in safe deployment of blocking irregular algorithms.

@@
